---
title: qsim
author: ~
date: '2017-06-15'
slug: qsim
categories: []
tags: []
---

```{r,message=FALSE}
knitr::opts_chunk$set(message=FALSE,comment='.')
library(rbenchmark)
library(mrgsolve)
library(dplyr)
```


`qsim` is a function in `mrgsolve` that is similar to `mrgsim`, but quicker.  In order to get that quickness, we lose some flexibility with the model.  But
in many circumstances, this is no big deal at all.  `qsim` works with a 
regular model object, an event object, and (optionally) an `idata_set`.
`qsim` returns simulations as a (plain/old) matrix so you will also
have to do a little more work on the back end.


Let's start with a benchmark.  For the example, load a 2-cmt model
and update the simulation end time.  We'll create a dosing event 
object (`ev`).  

```{r}
mod <- mread("pk2cmt", modlib()) %>% update(end=120)
ev <- ev(amt=100, ii=24,addl=2)
```

Finally, we will pass the dosing event object into `recmatrix` along
with the observation time grid (see `?stime`). 
```{r}
recs <- recmatrix(ev,stime(mod))
head(recs)
```

`recmatrix` takes
the dosing information in the event object, expands out all of 
the additional doses, schedules infusion end times (if `rate > 0`)
and inserts rows for observation records.  So the `recs` object
looks like a `data_set` but with slightly different information.  

Let's try the benchmark:
```{r}
benchmark(qsim(mod,recs),
          mrgsim(mod,events=ev),replications=300)[,c(1:4)]
```
```{r}
benchmark(qsim(mod,ev),
          mrgsim(mod,events=ev),replications=300)[,c(1:4)]
```

The idea
is to fix all of that __prior__ to running the simulation.  It won't always
give you a huge savings (for example, if every simulation has a different 
design or dose, you have to call `recmatrix` every time).  But it helps
with replicate simulation when you can build the design once beforehand,
and re-use for every replicate.

The time savings you get with `qsim` totally depend on the problem, including the duration of the simulation and the number of individuals.  
You might see a 100 speed up for simple simulation from a single set of parameters.  Or a 30% speed up for long simulations in a big population.  I would say you get the biggest speed up when it's the shortest, simplest problem. Using `qsim` in that case gets you out of all the overhead `mrgsolve` needs in order to do complicated simulations.  

But beware of the limitations here.  `mrgsolve` will call `recmatrix` on the event object even if you don't.  `qsim` always uses a single regimen for a single
fixed design with bioavability and infusion duration fixed prior to the problem.  



